


\documentclass[11pt]{article}
\usepackage{fullpage}
\usepackage{amsthm}
\usepackage{amsthm,amsmath,amsfonts,amssymb,amstext}
\usepackage{latexsym,ifthen,url,rotating}
\usepackage[usenames,dvipsnames]{color}


% --- -----------------------------------------------------------------
% --- Document-specific definitions.
% --- -----------------------------------------------------------------
\newtheorem{definition}{Definition}

\newcommand{\concat}{{\,\|\,}}
\newcommand{\bits}{\{0,1\}}
\newcommand{\Range}{{\mathrm{Range}}}
\newcommand{\A}{{\mathcal{A}}}

% --- -----------------------------------------------------------------
% --- The document starts here.
% --- -----------------------------------------------------------------
\begin{document}
%\maketitle
\sloppy

\noindent Kaitlin Poskaitis\\
Homework 6\\
Section 6\\

\begin{center}
\LARGE{\textbf{Homework 6}}\\
\large{\textbf{\emph{Due at the beginning of class on Wednesday, April 3}}}
\end{center}

\vspace{.1in}

\noindent\textbf{Instructions:} Point values for each problem are listed.
Write your solutions neatly or type them up.  Typed solutions will also be
accepted via Sakai.

\begin{enumerate}


\item (5 points) Let $X$ be a random variable on some sample space $S$ with
expected value $\mu$ and variance $\sigma^2$.  Let $X_1$ and
$X_2$ be \emph{identical independent copies} of $X$.  Find
$E((X_1 - X_2)^2)$ in terms of $\mu$ and $\sigma^2$.  Give some intuition for 
why your answer makes sense.

\item (3 points) Prove that $V(cX) = c^2 V(X)$ for any random variable $X$ and
real number $c$.
\begin{itemize}
	\item $V(cX)$
	\item $E[cX - E(cX)]^{2}$ By definition of variance
	\item $E[cX - cE(X)]^{2}$ By lemma $E(cX) = cE(x)$
	\item $E[c(X-E(X))]^{2}$ Algebra
	\item $c^{2}E[X-E(X)]^{2}$ Algebra
	\item $c^{2}V(X)$ By definition of variance
	\item QED
\end{itemize}

\item (5 points) Let $X$ be the random variable that is the sum of two
independent fair die rolls.  Let $Y$ be the outcome of the first roll minus the
outcome of the second roll.  Calculate $\mathrm{Cov}(X,Y)$.
Are $X$ and $Y$ correlated?
\begin{itemize}
	\item $E(X) = 7$
	\item $E(Y) = 0$
\end{itemize}

\item (4 points) Let $C$ be the random variable that is the number of heads in
$100$ independent fair coin flips.  What are $E(C)$ and $V(C)$?  Find upper
bounds on $P(C > 75)$, using Markov's Inequality and Tchebychev's Inequality.
\begin{itemize}
	\item $E(C) = 50$
	\item $E(C^{2}) = 75$
	\item $V(C) = \frac{1}{4}$
	\item Markov's:
	\begin{itemize}
		\item $P(C \geq 76) \leq \frac{50}{76}$
		\item $P(C \geq 76) \leq \frac{25}{38}$
	\end{itemize}
	\item Tchebychev's:
	\begin{itemize}
		\item $P(|C-\mu| \geq k\delta) \leq \frac{1}{k^{2}}$
		\item $|C-50| \geq \frac{k}{2}$
		\item $C-50 \geq \frac{k}{2}$
		\item $C \geq \frac{k}{2} + 50$
		\item $76 = \frac{k}{2} + 50$
		\item $k=13$
		\item $P(X \geq 76) \leq \frac{1}{169}$
	\end{itemize}
\end{itemize}

\item (4 points) Find an example where Markov's inequality is tight in the
follow sense: For each  positive integer $a$, find a non-negative random
variable $X$ such that $P(X\geq a) = E(X)/a$.
\begin{itemize}
	\item Let X be a random variable that can equal x or a.
	\item Define the probabilties as follows: $P(X=0) = 1-\frac{1}{a}$ and 
	$P(X=a) = \frac{1}{a}$
	\item $E(x) = 0+a\frac{1}{a} = 1$
	\item $P(X \geq a) = P(X = a) = \frac{1}{a} = \frac{E(X)}{a}$
\end{itemize}

\item (4 points) Prove the following: If $X$ is a non-negative random variable
with $E(X) = \mu$, then for every $k$, $P(X \geq k\mu) \leq 1/k$.
\textsf{Hint:  Apply Markov's inequality.}
\begin{itemize}
		\item $P(X \geq k\mu) \leq \frac{E(X)}{k\mu}$ By Markov's inequality
		\item $P(X \geq k\mu) \leq \frac{\mu}{k\mu}$ Value of $E(X)$
		\item $P(X \geq k\mu) \leq \frac{1}{k}$ Algebra
		\item QED
\end{itemize}

\item \textbf{Extra credit (4 points):} If $X$ is a non-negative
random variable with $E(X) = \mu$, then Markov's inequality tells us
that for every $a$, $P(X \geq a) \leq \mu/a$.  As we saw in class,
this bound is sometimes very loose.  In this problem we'll look at
a situation in which extra information can be used to tighten the
bound.

Suppose that we are told $X$ is \emph{bounded
from below by some number $b$}, meaning $P(X \geq b) = 1$.  Use this information
to find a tighter bound on $P(X \geq a)$.  \textsf{Hint:  Apply Markov's
inequality to the r.v.  $Y = X - b$.}

For concreteness, suppose $E(X) = 1000$ and $P(X \geq 500) = 1$.  Markov's
says that $P(X \geq 2000) \leq 1000/2000 = 1/2$.  Use your method find a
tighter bound on $P(X \geq 2000)$.

\end{enumerate}


\end{document}

