


\documentclass[11pt]{article}
\usepackage{fullpage}
\usepackage{amsthm}
\usepackage{amsthm,amsmath,amsfonts,amssymb,amstext}
\usepackage{latexsym,ifthen,url,rotating}
\usepackage[usenames,dvipsnames]{color}


% --- -----------------------------------------------------------------
% --- Document-specific definitions.
% --- -----------------------------------------------------------------
\newtheorem{definition}{Definition}

\newcommand{\concat}{{\,\|\,}}
\newcommand{\bits}{\{0,1\}}
\newcommand{\Range}{{\mathrm{Range}}}
\newcommand{\A}{{\mathcal{A}}}

% --- -----------------------------------------------------------------
% --- The document starts here.
% --- -----------------------------------------------------------------
\begin{document}
%\maketitle
\sloppy

\noindent Kaitlin Poskaitis\\
Section 1\\
Homework 7

\begin{center}
\LARGE{\textbf{Homework 7}}\\
\large{\textbf{\emph{Due at the beginning of class on Monday, April 15}}}
\end{center}

\vspace{.1in}

\noindent\textbf{Instructions:} Point values for each problem are listed.
Write your solutions neatly or type them up.  Typed solutions will also be
accepted via Sakai.

\begin{enumerate}


\item (1.5 points each) Find the generating function for the sequence
$a_0, a_1, a_2, \ldots$, where $a_k$ is each of the following.
Your solution does not need to be closed form.
\begin{enumerate}
\item $a_k = $ the number of solutions to $e_1 + e_2 + e_3 = k,$
where $0 \leq e_i \leq 4$ for each $i$.\\
$(1 + x + x^2 + x^3 + x^4)^3$
\item $a_k = $ the number of solutions to $e_1 + e_2 + e_3 + e_4 = k,$
where $0 \leq e_i < 4$ for each $i$, $e_1$ is odd, and $e_2$ is even.\\
$(1 + x^2 + x^4)^4$
\item $a_k = $ the number of solutions to $e_1 + e_2 + e_3 + e_4 = k,$
where $0 \leq e_i$ for each $i$.\\
$(1 + x + x^2 ... x^{\infty})^5$
\item $a_k = $ the number of solutions to $e_1 + e_2 + e_3 + e_4 + e_5= k,$
where $0 \leq e_i$ for each $i$, $e_1$ and $e_3$ are odd, and $e_2$
is even.\\
$(x + x^3 + x^5 ... + x^{\infty})^2(1 + x^2 + x^4 ... + x^{\infty})(1 + x +x^2 ... +
x^{\infty})^2$
\end{enumerate}

\item (3 points each) Model the following problems using a generation function,
which does not need to be in closed form:  
\begin{enumerate}
\item Count the number of outcomes of
rolling $6$ dice that sum to $r$.\\
$e_1 + e_2 + e_3 + e_4 + e_5 + e_6 = r$\\
$1 \leq e_i \leq 6$\\
$(x + x^2 + x^3 + x^4 + x^5 + x^6)^6$
\item Count the number of outcomes of rolling $6$ dice that sum to $r$,
where the first three dice are odd and the last three are even.
$e_1 + e_2 + e_3 + e_4 + e_5 + e_6 = r$\\
$1 \leq e_i \leq 6$\\
$e_1$ $e_2$ $e_3$ odd\\
$e_4$ $e_5$ $e_6$ even\\
$(x + x^3 + x^5)^3(x^2 + x^4 + x^6)^3$
\item Count the number of outcomes of rolling $6$ dice that sum to $r$,
where for each $i$ the $i$-th dice is not equal to $i$ (so
the first die is not $1$, the second is not $2$, and so on).
$e_1 + e_2 + e_3 + e_4 + e_5 + e_6 = r$\\
$1 \leq e_i \leq 6$\\
$e_i \neq i$\\
$(x^2 + x^3 + x^4 + x^5 +x^6)(x + x^3 + x^4 + x^5 + x^6)
(x + x^2 + x^4 + x^5 + x^6)(x + x^2 + x^3 + x^5 + x^6)
(x + x^2 + x^3 + x^4 + x^6)(x + x^2 + x^3 + x^4 + x^5)$

\end{enumerate}

\item (1.5 points each) Find the following coefficients.  Show your work.
\begin{enumerate}
\item The coefficient of $x^{10}$ in the series expansion of 
$(x^5+x^6 + x^7+ \cdots)^8$.\\
$0$
\item The coefficient of $x^{20}$ in the series expansion of 
$(x+x^2 + x^3+x^4+x^5) (x+x^2 + x^3+x^4+\cdots)^5 $.

Let $f(x) = (x + x^2 + x^3 + x^4 + x^5)$\\
Let $g(x) = (x + x^2 + \cdots)^5$\\
$f(x) = x(1 + x + x^2 + x^3 + x^4)$\\
$f(x) = x \frac{1-x^3}{1-x}$\\
$g(x) = [x(1 + x + x^2 + \cdots)]^5$\\
$g(x) = x^5(1 + x + x^2+ \cdots)^5$\\
$g(x) = x^5 \frac{1}{(1-x)^5}$\\
$f(x)g(x) = x^6 \frac{1}{(1-x)^6} (1 - x^3)$\\
Let $a(x) = x^6 \frac{1}{(1-x)^6}$\\
Let $b(x) = 1 - x^3$\\
In order to get $x^{20}$, need $a_{20}b_{0} + a_{17}b_{3}$\\
$a_{20} = {(14 + 6 - 1) \choose 14} = {19 \choose 14}$\\
$b_{0} = 1$
$a_{17} = {(11 + 6 -1) \choose 11} = {16 \choose 11}$\\
$b_{3} = -1$\\
Coefficient is ${19 \choose 14} - {16 \choose 11}$

\item The coefficient of $x^{12}$ in the series expansion of
$x^2/(1+x)^8$.
\item The coefficient of $x^{12}$ in the series expansion of
$1/(1+x^3)^2$.
\end{enumerate}

\end{enumerate}


\end{document}
